---
title: "WUP_04"
author: "Carmen Ramos-Caminos"
date: "Summer 2022"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
require(mosaic)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

```{r}
gssDataframe=read.csv("GSScsv.csv")
```

WORDSUM is a Numeric variable. DEGREE is a Categorical variable.  

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
boxplot(WORDSUM~DEGREE,data=gssDataframe)
```

From the box plot it appears that WORDSUM test scores are different. Individuals that have a high school degree or less education scored lower compared to individuals with higher education such as a graduate degree.    

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=gssDataframe)
```

It appears that there is a difference in the median score and ability on the WORDSUM test scores of graduate students versus students with high school or less education.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=gssDataframe)
summary.aov(model1)
```

I reject the null hypothesis $H_0$ based on the P-value is 0.000236
because the value is less than 0.05.


```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 15, 2, 2))
plot(TukeyHSD(model1),las=2)
```

Based on the graph there is a significant difference between high school vs graduate school taht does not include 0. There is also a difference between less then high school and graduate school since neither crosses 0. 
 
# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhaps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

We ae asked to find if there is a difference between the WORDSUM test scores of individuals varying from less than high education to graduate students. Based on the information and the P-value 0.000236, less than 0.05, I reject the null hypothosis and conclude that there is a difference between test scores of students with varying degrees.  